\textbf{Are regime changes always bad for markets? Evidence from daily financial data}

### 2021-08-11 doc/regime_changes_note.tex
Additional research note edits
Made additional edits to reformat paper to research note length.

\singlespacing
\begin{abstract}
\noindent
Political instability is commonly thought to discourage investment and reduce economic growth. We challenge this theory by finding that different types of instability have disparate effects on stock returns. Using an event study approach, we examine daily returns of national stock indices in every country that experienced an irregular regime change subject to data availability. Abnormal returns following resignations are large and positive (+4\%), while those following assassinations are negative and smaller in magnitude (-2\%). The impact of coups tends to be negative (-2\%), but examination of the 2002 failed coup against Hugo Chavez reveals returns of +10\%. We find suggestive evidence that authoritarian or anti-capitalist regime changes are more likely to be perceived negatively by investors than democratic or pro-business changes.  The expected direction and magnitude of abnormal returns is therefore dependent on the type of regime change and its expected impact on economic policy.

@@

Investors, multinational firms, development agencies, and aid organizations rank political risk as a top consideration when making investment decisions in emerging markets.\footnote{In the \citet{wipr2013} report (most recent), executives of multinational enterprises ranked political risk as the second most important constraint for foreign direct investment (FDI) in developing countries (after macroeconomic instability). In \citet{wipr2011} and  \citet{wipr2012}, political risk was ranked the most important constraint for FDI in developing countries - even greater than macroeconomic instability.}

Unstable governments have, for example, been argued to be less likely to invest in the legal system and the protection of property rights \citep{svensson1998investment}, and to be more likely to increases taxes \citep{devereux1998political}. This is supported by empirical research showing that political instability---as measured by regime change frequency, political violence, or indices of uncertainty derived from news reports---is negatively correlated with investment, financial development, and GDP growth in cross country regressions \citep{aisen2013does, alesina1996income, alesina1996political, fosu1992political, jong2009measurement, roe2011political, baker2016measuring} and that the variance of stock market returns increases in response to economic policy uncertainty \citep{leblang2005government, jensen2005market, liu2015economic}.

In contrast to these cross country studies, we: (1) examine the impact of various types of instability (coups, resignations, assassinations, and protests) separately, and (2) conduct our analysis using event studies of daily financial data. Event studies examine the effect of an unexpected event on stock prices \textit{at the time of the event} (i.e. a local average treatment effect) to determine if returns are systematically different from predicted. This interrupted time-series approach mitigates the endogeneity problems ubiquitous in previous cross country regressions. Confounding events would need to also occur on the same day, and do so for a large portion of all of our independently tested events in order to influence our estimates.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.,][]{fisman2001estimating,faccio2006politically,goldman2009politically}.}

@@

We analyze 13 successful coups, 8 assassinations, 15 forced resignations, and 11 public protests---the full sample of events for which national-level daily financial data is available.  Like previous research, we find that coups, assassinations, resignations, and public protests all cause large increases in financial volatility. However, we also find that abnormal returns are large and positive (+4\%) following resignations and tend to be negative and smaller in magnitude following assassinations and coups (-2\%). We also examine the failed 2002 Venezuelan coup, as it allows us to estimate the effects of both a pro-business coup (+10\%), and the reinstatement of a left-wing populist (-8\%) on the same market.

Our primary contributions are empirical and methodological. Empirically, we show that while political instability concerns investors, instability does not systematically depress investment. Investors recognize when instability may be in their favor. Methodologically, we employ a method less susceptible to endogeneity concerns, and integrate event studies with synthetical control methods to allow for control portfolios in situations where an obvious control portfolio candidate is not present. While we do not view theoretical development as our main contribution, we do find evidence that authoritarian regime changes are more likely to lead to negative returns, and that leaders who are clearly pro-business can be rewarded by investors, even if they use extra-judicial methods to take power.

@@

Conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}. However, if market returns reflect investor expectations about future economic growth and returns, such hypotheses may be too simplistic. Investors may view even a coup as a ``positive'' event if the current regime is anti-business or anti-global.

@@

Not all regime changes are equivalent. Resignations, assassinations, and coups have all been used as proxies for instability in previous research. However, some regime changes---such as the sudden resignation of an ineffective leader---may lead to increased stability. By contrast, assassinations can occur seemingly at random and a successor may be unclear. Coups can occur in the name of democracy or autocracy. These differences are often overlooked in past empirical strategies, which have proxied for instability in terms of the number of coups \citep{londregan1990poverty, alesina1996political}, assassinations or revolutions \citep{barro1991economic}, or have combined various metrics into single indices \citep{alesina1996income, venieris1986income, gupta1990economics, jong2009measurement}.\footnote{Note that these studies vary by metrics included in the indices, method of aggregation, and outcome variable of interest.} By contrast, we estimate effects separately for coups, assassinations, resignations, and protests.

An active debate also exists on the subject of ``good coups''---i.e., coups that lead to democratization or economic liberalization---and their effect on economic growth.  Most research suggests that ``good coups'' are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}, others argue that the negative effects of uncertainty dominate any positive effects \citep{alesina1996political}, or that the expected impact of coups  is neutral as some enhance growth while others depress it \citep{londregan1990poverty}. But while ``good coups'' may be the minority, they have become more frequent \citep{marinov2014coups} and could have positive economic effects \citep{meyersson2016political, girardi2018institution}.

Different types of regime changes may therefore have disparate effects on markets and/or economic growth. The theory and research above suggests that coups should on average lead to negative returns, but sometimes lead to positive returns when the coup's instigators are clearly more pro-market than the regime they replace. As our event study approach allows us to analyze individual coups separately from others, we can test this prediction. Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they typically signal the departure of an ineffective leader.

@@

Financial data are from the Global Financial Data database, which includes the longest available daily time series of stock prices. We collect national stock index data on every country in which there was a coup or coup attempt, an assassination or failed assassination, or a forced resignation and for which daily financial data is available.\footnote{The list of failed assassinations are from \citet{jones2009hit}. Coup attempts are those in category 2 in the CSP Coup d'etat dataset.} The longest available daily time series for these stock indices are listed in \autoref{tab:stock-list}.

Political data are primarily drawn from the Center for Systemic Peace's (CSP) Polity IV Coup d'etat dataset and Coup d'etat Events handbook. The Coup d'etat dataset includes the date of successful, attempted, plotted, and alleged coups. We focus on successful coups because the other classifications are contentious, and the choice of dataset and classification methodology would directly alter our findings.\footnote{\citet[p. 617]{needler1966political} has gone so far as to say that ``the categories of coups that were aborted, suppressed, or abandoned melt into each other and into a host of other non-coup phenomena so as to defy accounting,'' the Center for Systemic Peace states that it is ``confident that [its] list of successful coups is comprehensive'' but does not extend this confidence to attempted or failed coups, and \citet{powell2011global} state that it is ``difficult to identify more ambiguous forms of coup activity, such as coup failures, plots, and rumors.''}

The Coup d'etat Events handbook also provides a list of 1) auto-coups\footnote{Defined by Center for Systemic Peace as the ``occurrence of subversion of the constitutional order by a ruling (usually elected) executive and the imposition of an autocratic regime."}, 2) the ouster of leadership by foreign forces, 3) the ouster of leadership by rebel forces, 4) assassinations of the executive and 5) resignations of the executive due to poor performance and/or loss of authority. Daily financial data is available for countries in categories 4 and 5, so we supplement the coups with assassinations and resignations to form a list of ``irregular'' regime changes. The resignations are those in which the ruling executive was coerced to resign due to poor performance, public discontent and popular demonstrations.\footnote{Note that the Polity IV definition of ``poor performance'' is not synonymous with poor \textit{economic} performance, and in practice the reasons cited for resignation across events are: loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.}

@@

We further supplement the CSP data with leadership data from Archigos Version 4.1, which allows us to identify additional cases in which a ``leader lost power through irregular means.'' Irregular transfers of power are those in which leaders do not leave office ``in a manner prescribed by either explicit rules or established conventions.'' Nearly all removals by irregular means result from the threat or use of force (e.g. coups, revolts and assassinations).

A list of the political events in our dataset and whether they resulted in democratic or autocratic shifts is shown in \autoref{tab:regime-changes}. Coups tend to have the largest impact on the level of democratization as a number of countries have subsequently transitioned from democracies to anocracies or autocracies. On the other hand, neither assassinations nor resignations tend to have much impact on the level of democratization.

The discussion above shows that there is considerable debate about classification of regime changes. We recognize that some readers may feel certain events are missing based on their own substantive knowledge. We rely on commonly used third-party classifications in order to minimize the possibility that our results are driven by our own subjective classifications. The one exception is that we separately analyze the failed coup attempt in Venezuela in April 2002 in which President Hugo Chavez was removed from office for two days, as it provides a natural test of the impact of the seemingly successful removal of a left-wing populist with a pro-business regime, and the ensuing reinstatement of a left-wing populist.

\autoref{fig:AV-DR} plots the absolute value of daily stock returns averaged across all events and shows that the absolute value of returns on the event day (day 0) are significantly larger than on any other day. In addition, the magnitude of returns begins increasing just before the event day and remains high for a short period after. This suggests that financial volatility increases during the days surrounding regime changes, although it does not provide any evidence on mean returns. \nameref{sec: Impact of Political Instability on Stock Returns} tests these results more formally. It will first formally estimate the amount of volatility surrounding all irregular regime changes, then analyze coups, assassinations and resignations separately and determine what impact they have on mean returns.

\section{Estimation strategy} \label{sec: estimation}

@@

We first examine the effect of irregular regime changes on financial volatility. Our estimates are obtained from a generalized autoregressive conditional heteroskedasticity (GARCH) model estimated using 1000 pre-event days, the event day and 1000 post-event days. As in \citet{jensen2005market} and \citet{leblang2005government}, we use the GARCH (1,1) specification. In particular, for national stock index $i$,
\begin{align*}
R_{it}=\mu_i + \epsilon_{it},\hspace{1cm} \epsilon_{it}\sim \mathcal{N}\left(0,\sigma_{it}^2\right),
\end{align*}
where $\mu_i$ is a constant and,
\begin{align*}
\sigma_{it}^2&=\gamma_{i}+\alpha_{i}\epsilon_{i,t-1}^2+\beta_{i}\sigma_{i,t-1}^2.
\end{align*}

@@

For the magnitude and direction of the effect of irregular regime changes on stock returns, we follow the standard event study methodology as presented by, among others, \citet{mackinlay1997event} and \citet{campbell1997econometrics}. Normal performance is measured with a constant mean return model,
\begin{align} \label{eqn:market-model}
R_{it}=\mu_{i}+\epsilon_{it},
\end{align}
where $R_{it}$ is the logged return of national stock index $i$ on trading day $t$ and $\epsilon_{it}$ is the error term. We calculate abnormal returns (ARs), in an ``event window'' surrounding the date of each coup, $AR_{i\tau}=R_{i\tau}-\widehat{\mu}_i$, where $\tau$ is a date in the event window, and $\widehat{\mu}_i$ is estimated in an ``estimation window'' preceding the event window with \autoref{eqn:market-model}. We use a 41 day event window (i.e. 20 pre-event trading days, the event day, and 20 post-event trading days). The estimation window is the 250 trading days prior to the start of the event window. The abnormal returns are then used to generate cumulative abnormal returns (CARs) between event day $\tau_1$ and event day $\tau_2$: $CAR(\tau_1,\tau_2)=\sum_{\tau=\tau_1}^{\tau_2}AR_{i\tau}$. Standard errors and p-values are calculated using asymptotic t-statistics as in \citet{mackinlay1997event}.\footnote{It is appropriate to use the standard normal distribution to calculate test statistics because the length of the estimation window is sufficiently long (250 trading days).}

We define the event date as the first trading day in which the market could have reacted to news of the event.\footnote{For example, during the October 12, 1999 coup d'etat in Pakistan led by General Pervez Musharraf, the army announced that Prime Minister Nawaz Sharif had been dismissed after market hours at 10:15 pm. We code October 14th, the day in which the market re-opened, as the event day.} When events occurred on weekends, we change the event date to the following Monday. $(0,\tau-1)$ is used to denote the $\tau$-day period beginning with the event day and $(-1,\tau)$ to denote the negative $\tau$-day period beginning with the day prior to the event day.\footnote{In other words, for cumulative abnormal returns prior to the event date, we aggregate backwards starting at the day of the event. For example, $CAR(-1,-2$) is the sum of the abnormal returns on event date $-1$ and event day $-2$.} We present results for the sum of abnormal returns over the post-event windows of the event date only $(0,0$), the event date plus 6 days $(0,6$), and the event date plus 19 days $(0,19$). In addition, we present results for the pre-event windows $(-1,-7$) and $(-1,-20$).

We use a constant mean return model instead of a market model to maximize the number of observations and because our unit of analysis \textit{is} the country-wide market index (not firms).\footnote{Plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995, respectively.} To address concerns regarding use of a constant mean return model, we integrate synthetic control methodology \citep{abadie2003economic, abadie2010synthetic} with event study estimation. In short, we create a ``synthetic'' control portfolio for each event and compare observed returns to the synthetic returns. Each country is given a weight which represents its influence in the synthetic control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window.

More formally, let $\boldsymbol{R}_{k}$ be the vector of returns for the event country in the estimation window, $\boldsymbol{R}_{-k}$ be the vector of returns for all other countries in the estimation window, $\boldsymbol{X}_1=(\boldsymbol{R}_{k},\var(\boldsymbol{R}_{k}))$, $\boldsymbol{X}_0=(\boldsymbol{R}_{-k},\var(\boldsymbol{R}_{-k}))$, and $\boldsymbol{W}_{-k}$ be a $((N-1) \times 1)$ vector of weights where $N$ is the number of countries listed in \autoref{tab:stock-list}. Then $\boldsymbol{W}^*$ is chosen to minimize $(\boldsymbol{X}_1-\boldsymbol{X}_0\boldsymbol{W})'\boldsymbol{V}(\boldsymbol{X}_1-\boldsymbol{X}_0\boldsymbol{W})$ subject to $w_i\geq0$ $(i = 1,2,\ldots,N-1)$ and $\sum_i^{N-1} w_i = 1$, and the vector $\boldsymbol{V}$ is chosen so that stock returns for the control portfolio during the estimation window are are close as possible to the event country.\footnote{See \citet{abadie2003economic} for further details.} The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock-list}, and the results are presented in \nameref{subsec: Robustness}.

@@

\begin{figure}[!htb]
\includegraphics[scale = 0.8]{../figs/mean-volatility.pdf}
\caption{Mean of volatility estimates from GARCH(1,1) models}
\label{fig:volatility}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[max size={\textwidth}{.8\textheight}]{../figs/mean-car-by-regime-change-type.pdf}
\caption{Mean cumulative abnormal returns by type of regime change}\label{fig:mean-car-by-regime-change}
\end{figure}

\autoref{fig:mean-car-by-regime-change} shows the mean cumulative abnormal returns by each type of regime change we analyze.

Individual results for all coups in our sample and mean results for all coups can be found in \autoref{tab:AR-coups}. The average coup has a $-2.1\%$ event day AR. Event day ARs for the 1970 coup in Argentina, the 1991 coup in Thailand, the 1992 coup in Peru, and the 1999 coup in Pakistan are all negative and statistically different than zero. Moreover, all of these cases except Thailand have negative post-event CARs and pre-event CARs that are statistically indistinguishable from zero. In all of these cases, the coup either overthrew a democratically elected government or changed governance from one military ruler to another. The initial negative reaction followed by additional post-event negativity is consistent with the expected market reaction from a successful authoritarian coup followed by post-event consolidation of power.

@@

Like the majority of coups, there is evidence that assassinations decrease stock prices ( see \autoref{fig:mean-car-by-regime-change}). The results in \autoref{tab:AR-ass} are produced from analyses identical to those in \autoref{tab:AR-coups} but for assassinations rather than coups. The mean event day abnormal return is negative and statistically different than zero.

@@

There is no evidence of post or pre-event CARs in almost any of the assassinations. This is consistent with expectations as assassinations are typically not predictable. As with coups, the number of days that it took the stock market to rebound to pre-event levels is fairly low.\footnote{One exception is the assassination of William Mckinley in which the stock market didn't fully recover for 963 days, or almost 4 calendar years. However, this was likely caused by the Panic of 1901, which began when the stock market crashed on May 17th, 1901, and not by McKinley's death (although the assassination may have exacerbated the panic).}

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{fig:mean-car-by-regime-change} and \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\% and the positive returns are persistent and grow larger over time (mean 20-day CAR $\approx$ 12\%). Furthermore, event day ARs are only negative and statistically significant at even the ten percent level in two out of the fifteen resignations (Pakistan on April 19, 1993 and Tunisia on June 31, 2011).

These results are again consistent with our hypothesis that different events will have disparate effects, and that resignations may lead to positive returns as resignations typically occur because of poor performance and/or loss of authority. Among our sample of events, leaders were ousted following loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.

For example, the resignation of Ferdinand Marcos as President of the Philippines in February 1986 was associated with an approximately 13\% positive event day AR. Prior to his resignation, the Philippine regime was known for rampant corruption, crony capitalism, extreme inequality, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}. In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies \citep{overholt1986rise}.

@@

\section{Robustness} \label{subsec: Robustness}

There are some potential concerns with the results in the \nameref{subsec: Coups}, \nameref{subsec: Assassinations}, and \nameref{subsec: Resignations} sections. First, the abnormal returns could have been driven by factors unrelated to the regime changes.  Second, the true effects of regime changes on firm value may be underestimated if investors had apriori information. Third, the reported means are based on small sample sizes so confidence intervals based on normally distributed abnormal returns may be inappropriate.

>>>>

\singlespacing
\begin{abstract}
\noindent
Political instability is commonly thought to discourage investment and reduce economic growth.
We challenge this consensus by showing that instability does not systematically depress investment as investors recognize when instability may be in their favor. Using an event study approach, we examine daily returns of national stock indices in every country that experienced an irregular regime change subject to data availability. Returns following resignations are large and positive (+4\%), while those following assassinations are negative and smaller in magnitude (-2\%). The impact of coups tends to be negative (-2\%), but we show that a pro-capitalist coup results in large positive returns. We find suggestive evidence that authoritarian or anti-capitalist regime changes are more likely to be perceived negatively by investors than democratic or pro-business changes.  The impact of political instability on investment is therefore dependent on the type of regime change and its expected impact on economic policy.

@@

Investors rank political risk as a top consideration when investing in emerging markets.\footnote{Executives of multinational enterprises rank political risk as among the most important constraint for foreign direct investment (FDI) in developing countries \citep{wipr2011, wipr2012, wipr2013}.} Research shows that political instability is negatively correlated with investment, financial development, and GDP growth in cross country regressions \citep{aisen2013does, alesina1996income, alesina1996political, fosu1992political, jong2009measurement, roe2011political, baker2016measuring}, and that stock market variance increases in response to economic policy uncertainty \citep{leblang2005government, jensen2005market, liu2015economic}.

%Unstable governments have been argued to be less likely to invest in the legal system and the protection of property rights \citep{svensson1998investment}, and to be more likely to increases taxes \citep{devereux1998political}.

%---as measured by regime change frequency, political violence, or indices of uncertainty derived from news reports---

However, how different types of political instability affect markets remains unknown. In contrast to aforementioned cross country studies, we: (1) examine the impact of various types of instability (coups, resignations, assassinations, and protests) separately, and (2) conduct event studies of daily financial data. Event studies estimate a local average treatment effect of an unexpected event on stock prices \textit{at the time of the event}. This interrupted time-series approach mitigates the endogeneity problems in previous cross country regressions---confounding events would need to occur on the same day as instability, and do so for a large portion of all of our independently tested events in order to influence our estimates.

@@

We analyze the full sample of politically unstable events for which national-level daily financial data is available.\footnote{13 coups, 8 assassinations, 15 forced resignations, and 11 public protests.}  Like previous research, we find that all types of political instability cause large increases in financial volatility. However, we find that abnormal returns are large and positive (+4\%) following resignations and tend to be negative and smaller in magnitude following assassinations and coups (-2\%). We also examine the failed 2002 Venezuelan coup, as it allows us to estimate the effects of both a pro-business coup (+10\%), and the reinstatement of a left-wing populist (-8\%) on the same market.

Our primary contributions are empirical and methodological. Empirically, we provide the first estimates of financial market reactions disaggregated by type of political instability. We show that while political instability concerns investors, it does not systematically depress investment. Investors recognize when instability may be in their favor. The capital flows we document are not insubstantial---in many cases they represent larger shocks than the 2008 stock market crash. Methodologically, we (1) employ a method less susceptible to endogeneity concerns than previous studies, and (2) integrate synthetic control and event study methods to allow for control portfolios in situations where an obvious control portfolio is not present. We do not view theoretical development as our main contribution, but find evidence that authoritarian regime changes are more likely to lead to negative returns, and that leaders who are clearly pro-business can be rewarded by investors, even if they use extra-judicial methods to take power.

@@

Conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}. But if market returns reflect investor expectations about future economic growth and returns, such hypotheses may be too simplistic. Investors may view even a coup as a ``positive'' event if the current regime is anti-business or anti-global.

@@

Not all regime changes are equivalent. Some regime changes---e.g., the resignation of an ineffective leader---may foreshadow better policy. Assassinations can occur seemingly at random, and a successor may be unclear. Coups can be democratic or autocratic. These differences are often overlooked in past studies, which have proxied for instability using the number of coups \citep{londregan1990poverty, alesina1996political}, assassinations or revolutions \citep{barro1991economic}, or combined events into single indices \citep{alesina1996income, venieris1986income, gupta1990economics, jong2009measurement}.\footnote{Note that these studies vary by metrics included in the indices, method of aggregation, and outcome variable of interest.} By contrast, we estimate effects separately for coups, assassinations, resignations, and protests.

Debate also exists on the subject of ``good coups''---i.e., coups that lead to democratization or economic liberalization---and their effect on economic growth.  Most research suggests ``good coups'' are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}. Others argue the negative effects of uncertainty dominate any positive effects \citep{alesina1996political}, or that the expected impact of coups is neutral as some enhance growth while others depress it \citep{londregan1990poverty}. But while ``good coups'' may be the minority, they have become more frequent \citep{marinov2014coups} and could have positive economic effects \citep{meyersson2016political, girardi2018institution}.

Different types of regime changes may therefore have disparate effects on markets and/or economic growth. Previous research suggests that coups should on average lead to negative returns, but sometimes lead to positive returns when the coup's instigators are clearly more pro-market than the regime they replace.  Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they typically signal the departure of an ineffective leader.

@@

Financial data is from the Global Financial Data database---the longest available daily time series of stock prices. We collect national stock index data for every country in which there was a coup, assassination, or resignation and daily financial data is available.\footnote{The list of failed assassinations are from \citet{jones2009hit}. Coup attempts are those in category 2 in the CSP Coup d'etat dataset.} These stock indices are listed in \autoref{tab:stock-list}.

Political data are primarily drawn from the Center for Systemic Peace's (CSP) Polity IV Coup d'etat dataset and Coup d'etat Events handbook. We form a list of ``irregular'' regime changes from successful coups,\footnote{For example, \citet[p. 617]{needler1966political} states that ``the categories of coups that were aborted, suppressed, or abandoned melt into each other and into a host of other non-coup phenomena so as to defy accounting,'' the CSP is ``confident that [its] list of successful coups is comprehensive'' but does not extend this confidence to attempted or failed coups, and \citet{powell2011global} state that it is ``difficult to identify more ambiguous forms of coup activity, such as coup failures, plots, and rumors.''} assassinations of the executive, and resignations of the executive\footnote{Resignations are those in which the ruling executive was coerced to resign due to poor performance, public discontent and popular demonstrations.Note that the Polity IV definition of ``poor performance'' is not synonymous with poor \textit{economic} performance, and in practice the reasons cited for resignation across events are: loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.} as daily financial data is available for countries in these categories. We supplement the CSP data with Archigos Version 4.1 leadership data, which allows us to identify some additional cases of coups, assassinations, and protests. A list of the political events in our dataset and whether they resulted in democratic or autocratic shifts is shown in \autoref{tab:regime-changes}.

%in which a ``leader lost power through irregular means.'' Irregular transfers of power are those in which leaders do not leave office ``in a manner prescribed by either explicit rules or established conventions.'' Nearly all removals by irregular means result from the threat or use of force (e.g. coups, revolts and assassinations).

There is considerable debate about classification of regime changes. We recognize that some readers may feel certain events are missing. We rely on common third-party classifications to minimize the possibility our results are driven by our own classifications. The one exception is that we separately analyze the 2002 failed coup  in Venezuela, as it provides a natural test of the impact of the seemingly successful removal of a left-wing populist with a pro-business regime, and the ensuing reinstatement of a left-wing populist.

%in which President Hugo Chavez was removed from office for two days

%\footnote{Defined by CSP as the ``occurrence of subversion of the constitutional order by a ruling (usually elected) executive and the imposition of an autocratic regime."}

@@

%Coups tend to have the largest impact on the level of democratization as a number of countries have subsequently transitioned from democracies to anocracies or autocracies. On the other hand, neither assassinations nor resignations tend to have much impact on the level of democratization.

%\autoref{fig:AV-DR} plots the absolute value of daily stock returns averaged across all events. The returns on the event day (day 0) are significantly larger than on any other day. The magnitude of returns begins increasing just before the event day and remains high for a short period after. This suggests that financial volatility increases during the days surrounding regime changes, but says nothing about mean returns. \nameref{sec: Impact of Political Instability on Stock Returns} will first formally estimate the volatility surrounding all irregular regime changes, then analyze coups, assassinations and resignations separately and determine their impact on mean returns.

\section{Estimation} \label{sec: estimation}

@@

To examine the effect of irregular regime changes on financial volatility, we use a generalized autoregressive conditional heteroskedasticity (GARCH) model estimated using 1000 pre-event days, the event day and 1000 post-event days. As in \citet{jensen2005market} and \citet{leblang2005government}, we use the GARCH (1,1) specification. In particular, for national stock index $i$,
\begin{align*}
R_{it}=\mu_i + \epsilon_{it},\hspace{1cm} \epsilon_{it}\sim \mathcal{N}\left(0,\sigma_{it}^2\right),
\end{align*}
where $\mu_i$ is a constant and,
\begin{align*}
\sigma_{it}^2&=\gamma_{i}+\alpha_{i}\epsilon_{i,t-1}^2+\beta_{i}\sigma_{i,t-1}^2.
\end{align*}

@@

%\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.,][]{fisman2001estimating,faccio2006politically,goldman2009politically}.}

To estimate the magnitude and direction of the effect of irregular regime changes on stock returns, we follow standard event study methodology \citep{mackinlay1997event, campbell1997econometrics}. Normal performance is measured with a constant mean return model,
\begin{align} \label{eqn:market-model}
R_{it}=\mu_{i}+\epsilon_{it},
\end{align}
where $R_{it}$ is the logged return of national stock index $i$ on trading day $t$ and $\epsilon_{it}$ is the error term. We calculate abnormal returns (ARs) in an ``event window'' around the date of each event, $AR_{i\tau}=R_{i\tau}-\widehat{\mu}_i$, where $\tau$ is a date in the event window, and $\widehat{\mu}_i$ is estimated in an ``estimation window'' preceding the event window with \autoref{eqn:market-model}. We use a 41 trading day event window (20 pre-event days, event day, and 20 post-event days) and 250 trading day estimation window. The ARs are then used to calculate cumulative abnormal returns (CARs) between event day $\tau_1$ and event day $\tau_2$: $CAR(\tau_1,\tau_2)=\sum_{\tau=\tau_1}^{\tau_2}AR_{i\tau}$. Standard errors and p-values are calculated using asymptotic t-statistics as in \citet{mackinlay1997event}.\footnote{It is appropriate to use the standard normal distribution to calculate test statistics because the length of the estimation window is sufficiently long (250 trading days).}

The event date is the first trading day a market could react to news of the event. $(0,\tau-1)$ denotes the $\tau$-day period beginning with the event day and $(-1,\tau)$ denotes the negative $\tau$-day period beginning with the day prior to the event day. We present results for the sum of abnormal returns over the post-event windows of the event date only $(0,0$), the event date plus 6 days $(0,6$) and 19 days $(0,19$), and the pre-event windows $(-1,-7$) and $(-1,-20$).

%\footnote{In other words, for cumulative abnormal returns prior to the event date, we aggregate backwards starting at the day of the event. For example, $CAR(-1,-2$) is the sum of the abnormal returns on event date $-1$ and event day $-2$.}

%If an event occurred on a weekend, the event date is the following Monday.

We do not use a market model (where a market index is included as a control) to maximize the number of observations and because our unit of analysis \textit{is} the country-wide market index (not firms).\footnote{Plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995, respectively.} To address concerns regarding use of a constant mean return model, we combine synthetic control methods \citep{abadie2003economic, abadie2010synthetic} with event study estimation. We create a ``synthetic'' control portfolio for each event, where each country is given a weight representing its influence in the control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window.

Formally, let $\boldsymbol{R}_{k}$ be the vector of returns for the event country in the estimation window, $\boldsymbol{R}_{-k}$ be the vector of returns for all other countries in the estimation window, $\boldsymbol{X}_1=(\boldsymbol{R}_{k},\var(\boldsymbol{R}_{k}))$, $\boldsymbol{X}_0=(\boldsymbol{R}_{-k},\var(\boldsymbol{R}_{-k}))$, and $\boldsymbol{W}_{-k}$ be a $((N-1) \times 1)$ vector of weights where $N$ is the number of countries listed in \autoref{tab:stock-list}. Then $\boldsymbol{W}^*$ is chosen to minimize $(\boldsymbol{X}_1-\boldsymbol{X}_0\boldsymbol{W})'\boldsymbol{V}(\boldsymbol{X}_1-\boldsymbol{X}_0\boldsymbol{W})$ subject to $w_i\geq0$ $(i = 1,2,\ldots,N-1)$ and $\sum_i^{N-1} w_i = 1$, and the vector $\boldsymbol{V}$ is chosen so that stock returns for the control portfolio during the estimation window are are close as possible to the event country.\footnote{See \citet{abadie2003economic} for further details.} The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock-list}.

%and the results are presented in \nameref{subsec: Robustness}.

@@

\begin{figure}[!htb]
\includegraphics[scale = 0.75]{../figs/mean-volatility.pdf}
\caption{Mean of volatility estimates from GARCH(1,1) models}
\label{fig:volatility}
\end{figure}

@@

\autoref{fig:mean-car-by-regime-change} shows the mean CARs by each type of regime change we analyze for the event day, as well 20 days before and after the event. Individual results for all coups in our sample and mean results for all coups can be found in \autoref{tab:AR-coups}.

\begin{figure}[!htb]
\centering
\includegraphics[scale = 0.75]{../figs/mean-car-by-regime-change-type.pdf}
\caption{Mean cumulative abnormal returns by type of regime change}\label{fig:mean-car-by-regime-change}
\end{figure}

The average coup has a $-2.1\%$ event day AR. Event day ARs for the 1970 coup in Argentina, the 1991 coup in Thailand, the 1992 coup in Peru, and the 1999 coup in Pakistan are all negative and statistically different than zero. Moreover, all of these cases except Thailand have negative post-event CARs and pre-event CARs that are statistically indistinguishable from zero. In all of these cases, the coup either overthrew a democratically elected government or changed governance from one military ruler to another. The initial negative reaction followed by additional post-event negativity is consistent with the expected market reaction from a successful authoritarian coup followed by post-event consolidation of power.

@@

Like the majority of coups, we find that assassinations decrease stock prices (see \autoref{fig:mean-car-by-regime-change} and \autoref{tab:AR-coups}). The mean event day abnormal return is negative and statistically different than zero.

@@

There is no evidence of post or pre-event CARs in almost any of the assassinations. This is consistent with expectations as assassinations are typically not predictable. As with coups, the number of days that it took the stock market to rebound to pre-event levels is fairly low.\footnote{One exception is the assassination of William Mckinley in which the stock market didn't fully recover for 963 days. However, this was likely caused by the Panic of 1901, which began when the stock market crashed on May 17th, 1901, and not by McKinley's death.}

% (although the assassination may have exacerbated the panic)

@@

Unlike coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{fig:mean-car-by-regime-change} and \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\%, and the positive returns grow larger over time (mean 20-day CAR $\approx$ 12\%). Event day ARs are only negative and significant at even the ten percent level in two of the fifteen resignations.

These results are consistent with our hypothesis that resignations may lead to positive returns as they typically occur due to poor performance and/or loss of authority. Among our sample of events, leaders were ousted following loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.

%For example, the resignation of Ferdinand Marcos as President of the Philippines in February 1986 was associated with an approximately 13\% positive event day AR. Prior to his resignation, the Philippine regime was the least preferred site for foreign investment amongst East Asian capitalist economies due to rampant corruption, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}.

%In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies \citep{overholt1986rise}.

@@

\vspace{-0.05cm}
\section{Robustness} \label{subsec: Robustness}

Potential concerns with the results in the \nameref{subsec: Coups}, \nameref{subsec: Assassinations}, and \nameref{subsec: Resignations} sections are that: (1) the ARs could be driven by factors unrelated to the regime changes, (2) the effects of regime changes may be underestimated if investors had apriori information, and (3) confidence intervals based on normally distributed ARs may be inappropriate.
